---
title: "Data pre-process file"
output: html_document
date: "2023-11-07"
---

I have an excel file containing data from a metabolomics study. The data was taken from the paper "Women with temporomandibular disorders: Untargeted proton nuclear magnetic resonance spectroscopy-based metabolomics of saliva and psychological instruments dataset". Study authors: Sanches et al. 

There is only one excel file. 

Load the packages needed to preprocess the data:
```{r}
library(tidyverse)
library(readxl)
library(countrycode)
```
  
Read the Excel file as a dataframe:
```{r}
RAW_DATA_metabolomic_of_saliva <- read_excel("/Users/anaruxandrastanciu/test2/data-raw/RAW DATA metabolomic of saliva.xlsx")
```

Get a summary of the dataframe:
```{r}
summary(RAW_DATA_metabolomic_of_saliva)
```

The sources of these data are as follows:

- Sanches, M. L., Sforça, M. L., Turco, E. G. L., Faber, J., Smith, R. L., & Moraes, L. O. C. (2020). Women with temporomandibular disorders: Untargeted proton nuclear magnetic resonance spectroscopy-based metabolomics of saliva and psychological instruments dataset. Data in brief, 34, 106677. https://doi.org/10.1016/j.dib.2020.106677. Accessed Nov. 7th 2023

Save to data folder:
```{r}
usethis::use_data(RAW_DATA_metabolomic_of_saliva,
                  overwrite = TRUE)
```
